A Regularized Time Varying Vector Error Correction Model

### 2016-03-22 articles-conferences/elsevier-2016/journal.tex
Inicio articulo 2016

\title{A Time Varying Vector Error Correction Model Applied to Foreign Exchange
Market}
\author{Paola Arce \and Jonathan Antognini \and Luis Salinas \and Werner Kristjanpoller}

@@

In this study, we propose an online VECM which optimizes how model parameters
are obtained by using only relevant past data and RR instead of OLS.  Our
proposal also takes advantage of the long-run relationship between the time
series in order to obtain improved execution times. Our proposed method is
tested using synthetic data and foreign exchange rates.
Results show that our proposal enables execution time to be reduced by at least
50\% while maintaining good accuracy levels.
\end{abstract}

\input{src/intro}
\input{src/background}
\input{src/methodology}
\input{src/results}
\input{src/conclusions}
\bibliographystyle{apalike}
\bibliography{reference}
%\begin{thebibliography}{9}
%
%\bibitem{smola}
%Bernhard Scholkopf and Alexander J. Smola. 
%\emph{Learning with Kernels: Support Vector Machines, 
%Regularization, Optimization, and Beyond.}.
%MIT Press, Cambridge, MA, USA, 2001
%
%\bibitem{wiki}
%\url{http://en.wikipedia.org/wiki/Reproducing_kernel_Hilbert_space}
%
%\bibitem{wikibooksIsom}
%\url{http://en.wikibooks.org/wiki/Linear_Algebra/Definition_and_Examples_of_Isomorphisms}
%
%\bibitem{wolframREq}
%\url{http://mathworld.wolfram.com/EquivalenceRelation.html}
%
%\bibitem{wikiHomoM}
%\url{http://en.wikipedia.org/wiki/Homomorphism}
%
%
%\end{thebibliography}

>>>>

\title{A Regularized Time Varying Vector Error Correction Model}
\author{Paola Arce \and Jonathan Antognini \and Luis Salinas \and Werner Kristjanpoller}

@@

In this study, we propose an online VECM which optimizes how model parameters
are obtained by using only relevant past data and RR instead of OLS.  
\end{abstract}

%\input{src/intro}
\input{src/background}
%\input{src/methodology}
%\input{src/results}
%\input{src/conclusions}
\bibliographystyle{apalike}
\bibliography{reference}
%\begin{thebibliography}{9}
%
%\bibitem{smola}
%Bernhard Scholkopf and Alexander J. Smola. 
%\emph{Learning with Kernels: Support Vector Machines, 
%Regularization, Optimization, and Beyond.}.
%MIT Press, Cambridge, MA, USA, 2001
%
%\bibitem{wiki}
%\url{http://en.wikipedia.org/wiki/Reproducing_kernel_Hilbert_space}
%
%\bibitem{wikibooksIsom}
%\url{http://en.wikibooks.org/wiki/Linear_Algebra/Definition_and_Examples_of_Isomorphisms}
%
%\bibitem{wolframREq}
%\url{http://mathworld.wolfram.com/EquivalenceRelation.html}
%
%\bibitem{wikiHomoM}
%\url{http://en.wikipedia.org/wiki/Homomorphism}
%
%
%\end{thebibliography}

### 2016-03-22 articles-conferences/elsevier-2016/src/background.tex
Inicio articulo 2016

\subsection{Spurious Regression}

Regression analysis is widely used to test economic theories. But sometimes
regression is done between two variables, lets say $x_t$ and $y_t$, which are
not causally related. This occurs at the presence of a third stochastic variable
$z_t$ which is the true causal of the first two variables $x_t$ and $y_t$. The
regression measure R-squared gives high values in presence of spurious
relationships. This measure would means that the model explains all the
variability of the response data around its mean which is not true.

The R-squared measure is calculated as follows:

\begin{equation}
\label{eq:rsquared}
R^2 = 1 - \frac{\sum_{t=1}^T \epsilon_t^2}{\sum_{t=1}^T(y_t - \bar{y})^2}
\end{equation}

In presence of spurious relationships the denominator of equation
\ref{eq:rsquared} becomes very large because a large weight is placed on extreme
observations on either side of the mean $\bar{y}$.
In order to detect this, other statistic must be considered such as the
Durbin-Watson statistic.

% Acá agregar gráfico de spurious regression

\subsection{Integration and Cointegration}
A time series $\mathbf{y}$ is said to be integrated of order $d$ if after
differentiating the variable $d$ times, we get an I(0) process, more precisely:

\[
(1-L)^d \mathbf{y} \sim \text{I(0)} \, ,
\]

\noindent where I(0) is a stationary time series and $L$ is the lag operator:

\[
(1-L)\mathbf{y} = \Delta \mathbf{y}=\mathbf{y}_t  -\mathbf{y}_{t-1} \quad \forall t
\]

Let $\mathbf{y}_t = \{\mathbf{y}^1, \dots, \mathbf{y}^l\}$ be a set of $l$
stationary time series I(1) which are said to be cointegrated if a vector
$\beta=[\beta(1),\dots,\beta(l)]^\top \in \mathbb{R}^l$  exists such that the
time series,

\begin{equation}
 \mathbf{Z}_t:= \beta^\top \mathbf{y}_t = \beta(1) \mathbf{y}^1 + \dots + \beta(l) \mathbf{y}^l \sim
 \text{I(0)}\, .
\end{equation}

In other words, a set of I(1) variables is said to be cointegrated if
a linear combination of them exists which is I(0).

@@

\subsection{Unit Root Tests}
In order to use VECM the time series are required to be I(1) which is detected
using a unit root test. The Dickey-Fuller (ADF) test ~\cite{dickey1979} is a
unit root test where the null hypothesis is that the
variable has a unit root.

If we have a variable with the form:
\[
\mathbf{x}_t = \alpha \mathbf{x}_{t-1} + \epsilon_t
\]

\noindent and if we substract $\mathbf{x}_{t-1}$ from both sides we get:

\begin{equation}
\label{eq:DF}
\Delta \mathbf{x}_t = (\alpha-1) \mathbf{x}_{t-1} + \epsilon_t
\end{equation}

Dickey fuller test consist on determine if $\alpha$ equals 1. The alternative
hypothesis is $\alpha < 1$ which means that $\mathbf{x}_t$ is a stationary
process.

In order to include more than 1 lagged variable the Augmented Dickey-Fuller
test was developed and their tabulated critical values were obtained
by~\cite{mackinnon2010}. The number of lags is often selected using Schwartz
criterion.

\subsection {Cointegration Tests}

Johansen method~\cite{johansen1988} allows to find cointegration among several
I(1) time series. This test is more general than the
Engle-Granger~\cite{engle1987} test which only allow to find cointegration
between two time series.  Johansen procedure takes the form of a
likelihood-ratio test which determines $r$ cointegration relationships. The
method provides two tests based on an eigenvector decomposition of the matrix
$\Omega = \alpha \beta^\top$ on equation~\ref{eq:vec} called trace and maximal
eigenvalue statistic.  Once the number of cointegration number is determined,
$r$ more significative eigenvectors provided by Johansen procedure are used to
get the cointegration vector matrix $\beta^\top$.  Critical values for both
tests are limited to 12 and 5 variables in the trace and maximal eigenvalue
statistic respectively.

The trace test null hypothesis is $H_0: r = r_0$ against $H_1: r > r_0$ and the
maximum eigenvalue test null hypothesis is $H_0: r = r_0$ against $H_1: r =
r_0+1$ starting with $r_0 = 0$.  In the latest test, if we start with $r_0=0$
and $H_0$ is rejected $H_1: r = 1$ will be accepted which is not true if there
are more than one cointegration relationship, this is why this test is less
powerful and the trace test is commonly used.

@@

%Descomentar en la tesis
\textbf{Demo}\quad

Since $\mathbf{A}$ is singular the problem shown in
equation~(\ref{eq:regressionproblem}) has not solution, the minimum norm given
by equation~(\ref{eq:MP}) is obtained by solving the equivalent problem:

\begin{equation*}
\label{eq:proyectorsol}
\mathbf{A \hat{X} = PB} 
\end{equation*}

\noindent where $\mathbf{P=U_1 U_1^\top}$ is the projection onto the
Col($\mathbf{A}$).

Since $\mathbf{V} = [\underset{(n \times k)}{\mathbf{V_1}} |
\underset{(n \times k)}{\mathbf{V_2}}]$ and $\mathbf{V_1^\top V_2 =
0}$ we can express $\mathbf{\hat{X}} = \mathbf{V_1 x_1 + V_2 x_2}$
with $\mathbf{x_2=0}$ because $\mathbf{\hat{X}}$ lives in the
$\text{Row}(\mathbf{A})$ given by $\mathbf{V_1}$, so we have:

\begin{eqnarray*}
\mathbf{A \hat{X}} &=& \mathbf{PB} \\
\mathbf{U_1 \Sigma_1 V_1^\top \hat{X}} &=& \mathbf{U_1 U_1^\top B} \\
\mathbf{ V_1^\top \hat{X}} &=&  \mathbf{\Sigma_1^{-1} U_1^\top B} \\ 
\mathbf{ V_1^\top V_1 x_1} &=& \mathbf{\Sigma_1^{-1}
U_1^\top B} \\
\mathbf{x_1}&=& \mathbf{\Sigma_1^{-1} U_1^\top B}
\end{eqnarray*}

\noindent from this result we can obtain $\mathbf{\hat{X}}$ and
therefore the pseudo-inverse expression:

\begin{eqnarray*}
\mathbf{\hat{X}} &=& \mathbf{V_1 x_1} \\
                &=& \mathbf{V_1 \Sigma_1^{-1} U_1^\top B} \\
\mathbf{A^+} &=& \mathbf{V_1 \Sigma_1^{-1} U_1^\top} \, .
\end{eqnarray*}

@@

\textbf{Demo}\quad

To solve equation~(\ref{eq:optsolRR}) is equivalent to solve the following
optimization problem:

\begin{equation}
\label{eq:RRproblem2}
\underset{\mathbf{X}(\lambda)}{\text{min}} \quad \|
\mathbf{C}\mathbf{\mathbf{X}} - \mathbf{F} \|_2^2
\end{equation}

\begin{equation*}
	\mathbf{C}=\begin{bmatrix} \quad &\mathbf{A}& \quad \\ \hdashline \quad
& \lambda\mathbb{I}&\quad  \end{bmatrix} \quad \text{and} \quad
	\mathbf{F}=\begin{bmatrix} \quad &\mathbf{B}& \quad \\ \hdashline \quad
& \mathbf{0}&\quad  \end{bmatrix} \, .
\end{equation*}

Applying equation~(\ref{eq:MP}) and considering that $\mathbf{C}^\top
\mathbf{C} = \mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I}$ and 
$\mathbf{C}^\top \mathbf{F}=\mathbf{A}^\top \mathbf{B} $ we have:

\begin{eqnarray*}
\mathbf{X}(\lambda)&=&(\mathbf{C}^\top
\mathbf{C})^{-1}\mathbf{C}^\top \mathbf{F} \\
&=& (\mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I})^{-1} \mathbf{A}^\top \mathbf{B}
\end{eqnarray*}

@@

Figure~\ref{fig:shrinks} shows a visual example of the shrinking of the
coefficients:

\begin{figure}[h!]
\includegraphics[width=0.5\linewidth]{img/shrinks}
\caption{Shrink of regression coefficients}
\label{fig:shrinks}
\end{figure}

@@

\textbf{Demo}\quad

\begin{eqnarray*}
E[(B-\hat{f}(\mathbf{X}))^2] & =&
E[(B-f(\mathbf{X})+f(\mathbf{X})-\hat{f}(\mathbf{X}))^2] \\
&=& E[(B-f(\mathbf{X}))^2] +
E[(f(\mathbf{X})-\hat{f}(\mathbf{X}))^2] + \dots \\
& &
2\cancelto{0}{E[B-f(\mathbf{X})]}E[f(\mathbf{X})-\hat{f}(\mathbf{X})] \\
&=& \sigma^2 + MSE(\hat{f}(\mathbf{X}))
\end{eqnarray*}

\begin{eqnarray*}
    MSE(\hat{f}(\mathbf{X})) &=&
    E[f(\mathbf{X})-\hat{f}(\mathbf{X}))]^2 \\
    &=& E[f(\mathbf{X})-E[\hat{f}(\mathbf{X})] +
    E[\hat{f}(\mathbf{X})]-\hat{f}(\mathbf{X}))]^2 \\
    &=&  E[(f(\mathbf{X})-E[\hat{f}(\mathbf{X})])^2] +
      E[(\hat{f}(\mathbf{X})-E[\hat{f}(\mathbf{X})])^2] + \dots\\
    & & E[(f(\mathbf{X})-E[\hat{f}(\mathbf{X})])] 
    \cancelto{0}{E[(\hat{f}(\mathbf{X})-E[\hat{f}(\mathbf{X})])]} \\
    &=& Bias(\hat{f}(\mathbf{X}))^2 + Var(\hat{f}(\mathbf{X}))
\end{eqnarray*}

The bias of OLS is:

\begin{eqnarray*}
Bias(\hat{\mathbf{X}}) &=& E[\hat{\mathbf{X}}] - \mathbf{X} \\
&=& E[ (\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top \mathbf{B}] - \mathbf{X} \\
&=& E[ (\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top (\mathbf{AX})] - \mathbf{X}  \\
&=& \mathbf{X}  - \mathbf{X}  \\
&=&  0
\end{eqnarray*}

The bias of ridge regression when $\mathbf{A A^\top}$ is non-singular
can be obtained expressing ridge regression solution
$\mathbf{\lambda}$ in terms of OLS solution $\hat{\mathbf{X}}$:

\begin{eqnarray*}
\mathbf{X}(\lambda) &=&( \mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I})^{-1}\mathbf{B} \\
&=& (\mathbb{I} + \lambda (\mathbf{A}^\top \mathbf{A})^{-1})^{-1} (\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top \mathbf{B} \\
&=&  (\mathbb{I} + \lambda (\mathbf{A}^\top \mathbf{A})^{-1})^{-1}  \hat{\mathbf{X}} \\
&=& \mathbf{W} \hat{\mathbf{X}} 
\end{eqnarray*}

\noindent where $\mathbf{W}  = (\mathbb{I} + \lambda (\mathbf{A}^\top
\mathbf{A})^{-1})^{-1}  $ it is defined for simplicity. Ridge
regression bias is then obtained as:

\begin{eqnarray*}
Bias(\mathbf{X}(\lambda)) &=& E[\mathbf{X}(\lambda)] - \mathbf{X} \\
&=& E[\mathbf{W}\hat{\mathbf{X}}] - \mathbf{X} \\
&=&  \mathbf{W} \mathbf{X} - \mathbf{X} \neq 0 
\end{eqnarray*}

The variance of OLS is:

\begin{equation*}
Var(\hat{\mathbf{X}}) = \sigma^2 (\mathbf{A}^\top \mathbf{A} )^{-1}
\end{equation*}

\noindent and the variance of ridge regression is:

\begin{eqnarray*}
Var(\mathbf{X}(\lambda)) &=& Var(\mathbf{W}\hat{\mathbf{X}}) \\ 
&=& \mathbf{W}Var(\hat{\mathbf{X}})\mathbf{W}^\top \\
&=& \sigma^2 \mathbf{W}(\mathbf{A}^\top \mathbf{A} )^{-1}\mathbf{W}^\top
\end{eqnarray*}

\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{img/biasvariance}
\caption{Bias-variance tradeoff}
\label{fig:biasvariance}
\end{figure}

The figure~\ref{fig:biasvariance} shows the bias-variance tradeoff
given by equation~(\ref{eq:prederror}). Ridge regression shows an
increasing bias and a decreasing variance. Despite OLS has zero bias,
its variance is greater than ridge for small values of $\lambda$.
It can be shown that, in terms of prediction error, ridge (black line)
is lower than OLS (dotted line)~\cite{hoerl1970}.

>>>>

\subsection{Integration and Cointegration}\label{sec:coint}\  
Following Johansen \cite{johansen1995} we shall say that a stochastic process
$Y_t$ which satisfies $Y_t-E(Y_t) = \sum_{i=0}^\infty C_i\,\varepsilon_{t-i}$ is
called $I(0)$, and then we shall write $Y_t\sim I(0)$, whenever
$\sum_{i=0}^\infty C_i \neq 0$ and $\sum_{i=0}^\infty C_i\,z^i$ converges for
$z\in\mathbb{C}$ with $|z|<1$.  It is understood that the condition
$\varepsilon_t\sim iid(0,\sigma^2)$ holds.

A (vector) time series $\mathbf{y}_t$ is said to be {\em integrated of order\/}
$d$, and then we shall write $\mathbf{y}_t\sim I(d)$, whenever after $d$ times
(discrete) differentiation an stationary process is obtained~\cite{banerjee1993};
more precisely, whenever
$(1-L)^d\,\mathbf{y}_t\sim\text{I(0)}$, where $L$ is the usual lag operator:
$(1-L)\,\mathbf{y}_t = \Delta\mathbf{y}_t = \mathbf{y}_t-\mathbf{y}_{t-1}$ for
all $t$.

Note that this definition includes the scalar case as time series of
vectors of dimension 1; in this scalar case we will write the time series in
nonbold format.

Let $\mathbf{y}_t^\nu$, $\nu=1,\dots,l$, be a set of $l$ vector time series of
order $I(1)$.  They are said to be {\em cointegrated\/} if a vector
$\beta=[\beta(1),\dots,\beta(l)]^\top \in \mathbb{R}^p$ exists, such that the
time series,
\begin{equation}
\mathbf{Z}_t:= 
\sum_{\nu=1}^l \beta(\nu)\,\mathbf{y}_t^\nu\,\sim\,\text{I(0)}\,.
\end{equation}
In other words, a set of $I(1)$ time series is said to be cointegrated if a
linear combination of them exists, which is I(0).

@@

@@

@@

@@

@@

Ridge regression shows an
increasing bias and a decreasing variance. Despite OLS has zero bias,
its variance is greater than ridge for small values of $\lambda$.
It can be shown that, in terms of prediction error, ridge (black line)
is lower than OLS (dotted line)~\cite{hoerl1970}.

### 2016-03-22 articles-conferences/elsevier-2016/src/conclusions.tex
Inicio articulo 2016

\section{\uppercase{Conclusions}}
\label{sec:conclusions}
\noindent A new online vector error correction method was presented. 
We have shown that our proposed OVECM considerably reduces execution times
without compromising solution accuracy.  OVECM was compared with VECM and ARIMA
with the same sliding window sizes and OVECM outperformed both in terms of
execution time. Traditional VECM slightly outperformed our proposal but the
OVECM execution time is lower.
This reduction of execution time is mainly because OVECM avoids the cointegration vector
calculation using the Johansen method. The condition for getting new vectors is
given by the mean\_error variable which controls how many times the Johansen method is
called. Additionally, OVECM introduces matrix optimization in order to get the
new model in an iterative way.
We could see that our algorithm took much less than a minute at every
step. This means that it could also be used with higher frequency data and would
still provide responses before new data arrives.

For future study, it would be interesting to improve the out-of-sample forecast
by considering more explicative variables, to increase window sizes or trying new
conditions to obtain new cointegration vectors.

Since OVECM is an online algorithm which optimizes processing time, it could be
used by investors as an input for strategy robots. Moreover, some technical
analysis methods could be based on its output.

>>>>

\section{\uppercase{Conclusions}}
\label{sec:conclusions}
\noindent A new online vector error correction method was presented. The traditional VECM
was modified in order to use it in an online context. We have shown that our
proposed OVECM considerably reduces execution times without compromising
solution accuracy.  However the execution time of OVECM is only better than
SLVECM if we avoid the cointegration vector calculation using the Johansen
method, this is given by the avg\_error variable which controls how many times
the method is called.  We could see that our algorithm took much less than a
minute at every step. This means that it could also be used with higher
frequency data and would still provide responses before new data arrives. 
In all the presented scenarios, MAPE is still very high and for future
study, it would be interesting to improve the out-of-sample forecast by
considering more explicative variables, optimizing the windows size, the
number of lags or trying new condition to obtain new cointegration vectors.
Since OVECM is an online algorithm which optimizes processing time, it could be
used by investors as an input for strategy robots. Moreover, some technical
analysis methods could be based on its output.
